Modelling last glacial cycle ice dynamics in the Alps

### 2018-07-23 articles/cycle/response-monegato.tex

    Thank you for this detailed analysis. We want to point out that WorldClim
    is not a model, but is also based on observations. However we agree, that
    the regional dataset by \citep{Isotta.etal.2013}, tailored to the dense
    station network and steep topography of the Alps, may be a more robust
    reference precitation forcing that needs to be considered in future
    studies. A reference was added:

    \msquote{%
        Finally, modern precipitation data from WorldClim also bear
        uncertainties and exhibit local disagreement with other regional data
        \citep{Isotta.etal.2013}.}

    Besides, the following sentence was added to the discussion of glacier
    extent in the south-western Alps,

    \msquote{%
    However, geochronological data from the south-western Alps are sparse, and
    the LGM extent compilation \citep{Ehlers.Gibbard.2004} is inconsistent with
    more recent regional reconstructions \citep{Federici.etal.2017}.}

### 2018-07-23 articles/cycle/alps-cycle.tex
Address Giovanni's comment on precipitation.

    During MIS~2, all six simulations yield comparative modelled maximum
    extents (Fig.~\ref{fig:footprints}a). This result is inherent to the
    palaeo-climate forcing approach used, which involves a linear scaling of
    each palaeo-temperature anomaly record to a geomorphological
    reconstruction of the area glaciated by the Rhine Glacier piedmont lobe
    (Sect.~\ref{sec:paltemp}; Fig.~\ref{fig:footprints}a, black rectangle).
    Outside this benchmark area, all simulations underestimate ice extent in
    the western part of the model domain, and overestimate it in the eastern
    part, relative to the geomorphological ice limits
    (Fig.~\ref{fig:footprints}a). This result might indicate that the LGM
    temperature depression, here taken as homogeneous, was actually lower in
    the Eastern Alps than in the Western Alps, as previously shown by positive
    degree-day modelling of the central European palaeo-ice caps
    \citep{Heyman.etal.2013}. Alternatively, the east-west gradient in winter
    precipitation existing today (Fig.~\ref{fig:inputs}h) was
    perhaps enhanced during the LGM, as indicated by pollen reconstructions
    \citep{Wu.etal.2007}. Finally, the LGM extent of Alpine glaciers might have
    been affected by an east-west gradient in variables not accounted for
    by the PDD model, such as cloud cover or dust deposition.

@@

    On a more regional level, the maximum extent of the Adda, Oglio,
    Adige and Piave Glaciers in the central southern Alps is modelled
    several kilometres within the mapped LGM moraines (Fig.~\ref{fig:lgmvel}a).
    This appears to result
    from the palaeo-precipitation reduction used to force the model
    (Fig.~\ref{fig:footprints}b), which is likely unrealistic in at least this
    part of the model domain. On the other hand, the Durance glacier in the
    south-western Alps is modelled to extend several kilometers beyond the
    mapped limit (Fig.~\ref{fig:lgmvel}a), indicating that the LGM temperature
    depression was likely
    dampened by Mediterranean climate in this part of the model domain.

>>>>

    During MIS~2, all six simulations yield comparative modelled maximum
    extents (Fig.~\ref{fig:footprints}a). This result is inherent to the
    palaeo-climate forcing approach used, which involves a linear scaling of
    each palaeo-temperature anomaly record to a geomorphological
    reconstruction of the area glaciated by the Rhine Glacier piedmont lobe
    (Sect.~\ref{sec:paltemp}; Fig.~\ref{fig:footprints}a, black rectangle).
    Outside this benchmark area, all simulations underestimate ice extent in
    the western part of the model domain, and overestimate it in the eastern
    part, relative to the geomorphological ice limits
    (Fig.~\ref{fig:footprints}a). This result might indicate that the LGM
    temperature depression, here taken as homogeneous, was actually lower in
    the Eastern Alps than in the Western Alps, as previously shown by positive
    degree-day modelling of the central European palaeo-ice caps
    \citep{Heyman.etal.2013}. Alternatively, the east-west gradient in winter
    precipitation existing today (Fig.~\ref{fig:inputs}h) was
    perhaps enhanced during the LGM, as indicated by pollen reconstructions
    \citep{Wu.etal.2007}. The LGM extent of Alpine glaciers might have
    been affected by an east-west gradient in variables not accounted for
    by the PDD model, such as cloud cover or dust deposition. Finally, modern
    precipitation data from WorldClim also bear uncertainties and exhibit local
    disagreement with other regional data \citep{Isotta.etal.2013}.

@@

    On a more regional level, the maximum extent of the Adda, Oglio,
    Adige and Piave Glaciers in the central southern Alps is modelled
    several kilometres within the mapped LGM moraines (Fig.~\ref{fig:lgmvel}a).
    This appears to result
    from the palaeo-precipitation reduction used to force the model
    (Fig.~\ref{fig:footprints}b), which is likely unrealistic in at least this
    part of the model domain. On the other hand, the Durance glacier in the
    south-western Alps is modelled to extend several kilometers beyond the
    mapped limit (Fig.~\ref{fig:lgmvel}a), indicating that the LGM temperature
    depression was likely
    dampened by Mediterranean climate in this part of the model domain.
    However, geochronological data from the south-western Alps are sparse, and
    the LGM extent compilation \citep{Ehlers.Gibbard.2004} is inconsistent with
    more recent regional reconstructions \citep{Federici.etal.2017}.
